### 2022-05-05 paper/ase22/approach.tex
background
modify appraoch

\section{Specification- and Syntax-based Linguistic Capability Testing}
%% \label{sec:approach}

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figs/overview.pdf}
  \caption{\OverviewFigCaption}
\end{figure*}

\sw{Some paragraphs in this overview part should go earlier in the
  paper. I am just writing them here for now as I don't think we have
  them in the intro/background yet.}  We design and implement
\emph{Specification- and Syntax-based Linguistic Capability Testing
  (\tool{})} to automatically generate test cases to test the
robustness of sentiment analysis models. We identify four goals for a
large and effective test suite:

\begin{description}
\item[{\bf G1}] the test suite should contain realistic \sents;
\item[{\bf G2}] the test suite should cover diverse syntactic structures;
  \item[{\bf G3}] each test case should be
categorized into a \lc;
\item[{\bf G4}] the label of each test case should be
automatically and accurately defined.
\end{description}

\Cklst's templates generate complete and realistic \sents, and each
template maps to a \lc, satisfying {\bf G1} and {\bf
  G3}. But \Cklst only uses \sw{X} manually created templates to
generate its test suite; all test cases generated by the same template
share the same syntactic structure, thus violating {\bf G2}. In
addition, the label of each \Cklst test case has to be decided
manually, associated with each template, violating {\bf G4}.

We present \tool{}, a new \lc test case generation
tool, that satisfies all of these criteria.  \sw{I could not summarize
  a cohesive idea that drives our design. Leaving it here to fill in.
  Also, I felt my writing below still lacks justification for some
  design choices (e.g., why we do the differentiation).}
Figure~\ref{fig:overview} shows the overview of \tool{}, which
consists of two phases.  The \emph{specification-based seed
  generation} phase performs rule-based searches from a real-world
dataset ({\bf G1}) and template-based transformation to obtain the
initial seed \sents.  The search rules (e.g., search for \neu
\sents that do not include any \pstv or \ngtv words) and
transformation templates (e.g., \sw{add an example} \jl{\sent
  negation}) are defined in the \emph{\lc
  specifications}, which guarantee that each resulting seed conforms
to a specific \lc ({\bf G3}) and is labelled
correctly ({\bf G4}).

The \emph{syntax-based \sent expansion} phase expands the seed
\sents with additional syntactic elements (i.e., words
\sw{more?}\jl{and production ruls in \cfg}) to cover many real-world
syntactic structures ({\bf G2}). It first performs a syntax analysis
to identify the \pos (PoS) tags that can be inserted to each
seed, by comparing the PoS parse trees between the seed \sent and
many other \sents from a large reference dataset. Each identified
tag is inserted into the seed as a \emph{mask}. It then uses an NLP
recommendation model (i.e., BERT \cite{}) to suggest possible
words. If a resulting \sent is validated to be consistent with the
specification which additionally defines the rules for expansion
(e.g., the expanded word should be \neu), {\bf G3} and {\bf G4} are
still satisfied.  Last, because some validated \sents may include
unacceptable suggested words given the context \sw{is this the
  right motivation to do selection?} \jl{I edited the \sent. please
  let me know if it is clear}, we use a heuristic (i.e., the
confidence score from the NLP recommendation model) to select the more
realistic context-aware expanded \sents into \tool{}'s test suite.

@@

\subsection{Specification-based Seed Generation}
The seed generation phase of \tool starts by searching \sents in a
real-world dataset that match the rules defined in the \lc
specification, and then transforming the matched \sents using
templates to generate seed \sents that conform to individual
\lcs. The reasons for this design choice are twofold.  First, while
generally judging which \lc any \sent falls into and which label it
should have is infeasible, there exist simple rules and templates to
allow classifying the resulting \sents into individual \lcs and
with the correct labels, with high confidence.  This enables us to
test each \lc individually.  Second, searching from a real-world
dataset ensures that the \sents used as test cases for testing \lcs
are realistic and diverse. The diverse test cases are more likely to
achieve a high coverage of the target model's functionality in each
\lc, thus detecting more errors.
%
In this phase, \tool first search and selects \sents applicable to
the \lc in a given real-world dataset with search rules. In case that
the search rules only fulfill portion of the \lc specifications, the
selected \sents are not yet appropriate to become seed, we
transform the selected \sents into seed \sents using the
heuristic templates.
%
Table~\ref{tab:specification} shows the search rules and the
transformation templates of all 11 linguistic capabilities we
implemented in \tool. The first column shows the \lc type and its
description, and the second column shows the search rule and
transformation template used in each \lc. For LC1 and LC2, the NLP
models are evaluated in the scope of short \sents with selective
sentiment words. It does not require any guidance for its
transformation because the search rule alone is sufficient to conform
to the \lcs. On the other hand, search rules of LC3 and LC4 are not is
not enough to match their \lc specification, thus \tool uses heuristic
templates to conform the found \sents to the \lc. For example, the
selected \sents becomes seeds by preturbing them with the templates
and by negating the selected demonstrative \sents to conform to LC3
and LC4 respectively (see the third and forth rows in
Table~\ref{tab:specification}).
\sw{@Jaeseong: revise the rest of 3.1 based on
  Table~\ref{tab:specification}. I commented out the old text but it
  is still in the tex file.} \jl{I revised and added the explanation}

@@

The simple search rules and transformation templates used to generate
the seed \sents may limit the syntactic structures these seeds may
cover. To address this limitation, the syntax-based \sent expansion
phase extends the seed \sents to cover syntactic structures
commonly used in real-life \sents. Our idea is to differentiate the
parse trees between the seed \sents and the reference \sents
from a large real-world dataset. The extra PoS tags in the reference
parse trees are identified as potential syntactic elements for
expansion and inserted into the seed \sents as masks. We then use
masked language model to suggest the fill-ins. If the resulting
\sents still conform to the \lc specification,
they are added to \tool{}'s test suite. \sw{May have some redundancy
  and inconsistency with the overview part.}

@@

For each production of in each seed's parse tree (lines 3 and 4), we
extract its non-terminal at the left-hand-side (line 5), $s\_lhs$, and
the grammar symbols at the right-hand-side (line 6), $s\_rhs$. In line
7, the algorithm iterates through all productions in the reference
context-free grammar and match these that have the same non-terminal
at the left-hand-side as $s\_lhs$.  The right-hand-side of each
matched production is called $r\_rhs$.  If $s\_rhs$ consists of a
subset of the grammar symbols in $r\_rhs$ (line 8), the additional
symbols in the $r\_rhs$ are inserted as masks in the parse tree of
seed \sent, in their respective positions in the expanded production.
The left to right traversal of the leaves of an expanded parse tree
forms a masked \sent.  Lastly, due to the inefficient cost of
accessing full list of the masked \sents, we randomly select $k$
masked \sents for the next \sent expansion and validation phase when
the masked \sents are more than maximum number of masked \sents. The
random sampling is unbiased approach since it gives same chance to be
chosen. Thus, the random sample becomes representative of the
population of the masked \sents, and it efficiently shows the
usefulness of the \tool.  \sw{Add justification: why we need to select
  k masked \sents (performance?)  and why random makes sense.} \jl{I
  added the statement for it}

@@

In this phase, the words to fill in the masks in the masked \sents are
suggested by the BERT pretrained model~\cite{devlin2019bert}.  The
BERT is a transformer-based \nl model. It is pretrained on two tasks
of masked token prediction and next sentence prediction. As a result
of the training process, the BERT model suggests word for the mask
token according to its surrounding context in \sent. For each masked
token, multiple words are suggested ranked by their confidence scores.
\sw{Algorithm 1 does not require we only have one mask in the masked
  \sent. Can the model suggest multiple words at the same time?}
\jl{yes. it can predict multiple masked tokens at the same time.}
\sw{Say a bit more about the BERT model suggestion. E.g., it may
  suggest multiple words for the same mask but they are ranked?}
\jl{I added the explanation of BERT}
%
Because BERT model is not aware of the \lc
specification and the grammar symbol in the expanded parse tree, an
expanded \sent using the suggested words may no longer satisfy the
\lc specification. Therefore, we perform validation
on the suggested words and only accept them if the following three
criteria are met.

First, the PoS tag of the suggested word must match the PoS tag of the
expanded symbol in the parse tree. For the example in
Figure~\ref{fig:ExpEx}, the masked symbol is a ``NNS" (i.e., plural
noun); thus, the suggested word must also be a ``NNS". In this work,
we use \spacy, a free open-source library for \nlp, for extracting PoS
tags for each suggested word.\sw{Say how we obtain the PoS tag of a
  suggested word.}\jl{I added it} Second, it is required that the
sentiment of the expanded \sent becomes the same as the seed \sent. To
ensure this, the suggested words must be \neu.  \sw{Should we present
  this as part of specification, called expansion rule?} \jl{I did it
  because I think that there is no issue on mentioning it and that is
  how we assumed and did} Third, we additionally verify that the
expanded \sents satisfied the same search rules for the seed
\sent. Our goal for generating the expanded \sents is to use them for
evaluating the \sa models on the associated \lc in addition to the
seed \sent. It is only achieved when the expanded \sents are also met
with the same search rules for the \lc. For LC1 as an example, the
expanded \sent must still conform to the specification of the seed's
\lc specification. Therefore, the expanded \sents are required to be
short and to only have \neu adjectives and \nns. \sw{Say why we use
  the third criteria only for LC1 and LC2.}  \jl{I added the
  explanation}

\paragraph{Running example.} The third step in Figure~\ref{fig:ExpEx} shows the words suggested
by BERT. For this masked \sent, BERT suggested six words. Each word
is associated with the confidence score provided by BERT, the PoS tag,
and the sentiment. Among the six words, only ``ways'' and ``things''
are validated by \tool{} because they have the Pos tag ``NNS'' and are
\neu. In addition, it is found that both \sents meets the search
rule of the associated \lc of \SareqExOne. In the end, two \sents
of ``Or both ways'' and ``Or both things'' are generated.  \sw{Do we
  also check if the expanded \sent meets the
  specification?}\jl{Yes. I added the explanation of validation of \lc
  \req}

@@

\subsubsection{\Sent Selection}
\sw{@Jaeseong: Add how we select expanded \sents and motivate why.}
After 
\paragraph{Running example.} \sw{Refer to the example to say how we
  use the BERT score to select.} \jl{I dont think we need this
  subsection of \sent selection it is already explained at the
  previous stage with running example.}

>>>>

\section{Specification- and Syntax-based Linguistic Capability Testing}
\label{sec:approach}

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{figs/overview.pdf}
  \caption{\OverviewFigCaption}
\end{figure*}

% \sw{Some paragraphs in this overview part should go earlier in the
%   paper. I am just writing them here for now as I don't think we have
%   them in the intro/background yet.}

We design and implement a new NLP model testing method,
\emph{Specification- and Syntax-based Linguistic Capability Testing
  (\tool{})}, that automatically generate test cases with oracles to
test the robustness of sentiment analysis models. \tool{} addresses
all three challenges discussed above.

% We identify four goals for a
% large and effective test suite:

% \begin{description}
% \item[{\bf G1}] the test suite should contain realistic \sents;
% \item[{\bf G2}] the test suite should cover diverse syntactic structures;
%   \item[{\bf G3}] each test case should be
% categorized into a \lc;
% \item[{\bf G4}] the label of each test case should be
% automatically and accurately defined.
% \end{description}

% \Cklst's templates generate complete and realistic \sents, and each
% template maps to a \lc, satisfying {\bf G1} and {\bf
%   G3}. But \Cklst only uses \sw{X} manually created templates to
% generate its test suite; all test cases generated by the same template
% share the same syntactic structure, thus violating {\bf G2}. In
% addition, the label of each \Cklst test case has to be decided
% manually, associated with each template, violating {\bf G4}.

% We present \tool{}, a new \lc test case generation
% tool, that satisfies all of these criteria.  \sw{I could not summarize
%   a cohesive idea that drives our design. Leaving it here to fill in.
%   Also, I felt my writing below still lacks justification for some
%   design choices (e.g., why we do the differentiation).}
Figure~\ref{fig:overview} shows the overview of \tool{}, which
consists of two phases.  The \emph{specification-based seed
  generation} phase performs rule-based searches from a real-world
dataset and template-based transformation to obtain the initial seed
\sents.  The search rules (e.g., search for \neu \sents that do not
include any \pstv or \ngtv words) and transformation templates (e.g.,
negating a sentence) are defined in the \emph{\lc specifications},
which guarantee that each resulting seed conforms to a specific \lc
({\bf C2}) and is labelled correctly ({\bf C3}).

The \emph{syntax-based \sent expansion} phase expands the seed \sents
with additional syntactic elements (i.e., words) to cover many
real-world syntactic structures ({\bf C1}). It first performs a syntax
analysis to identify the \pos (PoS) tags that can be inserted to each
seed, by comparing the PoS parse trees between the seed \sent and many
other \sents from a large reference dataset. Each identified tag is
inserted into the seed as a \emph{mask}. It then uses an NLP
recommendation model (i.e., BERT \cite{}) to suggest possible
words. If a resulting \sent is validated to be consistent with the
specification which additionally defines the rules for expansion
(e.g., the expanded word should be \neu), {\bf C2} and {\bf C3} are
still satisfied. Last, because some validated \sents may include
unacceptable suggested words given the context, we use a heuristic
(i.e., the confidence score from the NLP recommendation model) to
select the more realistic context-aware expanded \sents into \tool{}'s
test suite.

@@

\subsection{Specification-based Seed Generation}
The seed generation phase of \tool starts by searching \sents in a
real-world dataset that match the rules defined in the \lc
specification, and then transforming the matched \sents using
templates to generate seed \sents that conform to individual \lcs. The
reasons for this design choice are twofold.  First, while generally
judging which \lc any \sent falls into and which label it should have
is infeasible, there exist simple rules and templates to allow
classifying the resulting \sents into individual \lcs and with the
correct labels, with high confidence.  This enables us to test each
\lc individually.  Second, searching from a real-world dataset ensures
that the \sents used as test cases for testing \lcs are realistic and
diverse. The diverse test cases are more likely to achieve a high
coverage of the target model's functionality in each \lc, thus
detecting more errors.
%
In this phase, \tool first search and selects \sents applicable to the
\lc in a given real-world dataset with search rules. In case that the
search rules only fulfill portion of the \lc specifications, the
selected \sents are not yet appropriate to become seed, we transform
the selected \sents into seed \sents using the heuristic templates.
%
Table~\ref{tab:specification} shows the search rules and the
transformation templates of all 11 linguistic capabilities we
implemented in \tool. The first column shows the \lc type and its
description, and the second column shows the search rule and
transformation template used in each \lc. For LC1 and LC2, the NLP
models are evaluated in the scope of short \sents with selective
sentiment words. It does not require any transformation because the
search rule alone is sufficient to conform to the \lcs. On the other
hand, search rules of LC3 to LC11 are not enough to match their \lc
specification, thus \tool uses heuristic templates to conform the
searched \sents to the \lc. For example, in LC3's transformation
template, the searched \sents become seeds by preturbing them with the
defined templates.  In LC4's transformation template, the searched
demonstrative sentences are negated.
% \sw{@Jaeseong: revise the rest of 3.1 based on
%   Table~\ref{tab:specification}. I commented out the old text but it
%   is still in the tex file.} \jl{I revised and added the explanation}

@@

The simple search rules and transformation templates used to generate
the seed \sents may limit the syntactic structures these seeds may
cover. To address this limitation, the syntax-based \sent expansion
phase extends the seed \sents to cover syntactic structures commonly
used in real-life \sents. Our idea is to differentiate the parse trees
between the seed \sents and the reference \sents from a large
real-world dataset. The extra PoS tags in the reference parse trees
are identified as potential syntactic elements for expansion and
inserted into the seed \sents as masks. We then use masked language
model to suggest the fill-ins. If the resulting \sents still conform
to the \lc specification, they are added to \tool{}'s test suite.

@@

For each production of in each seed's parse tree (lines 3 and 4), we
extract its non-terminal at the left-hand-side (line 5), $s\_lhs$, and
the grammar symbols at the right-hand-side (line 6), $s\_rhs$. In line
7, the algorithm iterates through all productions in the reference
context-free grammar and match these that have the same non-terminal
at the left-hand-side as $s\_lhs$.  The right-hand-side of each
matched production is called $r\_rhs$.  If $s\_rhs$ consists of a
subset of the grammar symbols in $r\_rhs$ (line 8), the additional
symbols in the $r\_rhs$ are inserted as masks in the parse tree of
seed \sent, in their respective positions in the expanded production.
The left to right traversal of the leaves of an expanded parse tree
forms a masked \sent.  Lastly, due to the inefficient cost of
accessing full list of the masked \sents, we randomly select $k$
masked \sents for the next \sent expansion and validation phase when
the masked \sents are more than maximum number of masked \sents. The
random sampling is unbiased approach since it gives same chance to be
chosen. Thus, the random sample becomes representative of the
population of the masked \sents, and it efficiently shows the
usefulness of the \tool.

@@

In this phase, the words to fill in the masks in the masked \sents are
suggested by the BERT pretrained model~\cite{devlin2019bert}.  The
BERT is a transformer-based \nl model. It is pretrained on two tasks
of masked token prediction and next sentence prediction. As a result
of the training process, the BERT model suggests word for the mask
token according to its surrounding context in \sent. For each masked
token, multiple words are suggested ranked by their confidence scores.
% \sw{Algorithm 1 does not require we only have one mask in the masked
%   \sent. Can the model suggest multiple words at the same time?}
% \jl{yes. it can predict multiple masked tokens at the same time.}
% \jl{I added the explanation of BERT}
%
Because BERT model is not aware of the \lc specification and the
grammar symbol in the expanded parse tree, an expanded \sent using the
suggested words may no longer satisfy the \lc
specification. Therefore, we perform validation on the suggested words
and only accept them if the following three criteria are met.

First, the PoS tag of the suggested word must match the PoS tag of the
expanded symbol in the parse tree. For the example in
Figure~\ref{fig:ExpEx}, the masked symbol is a ``NNS" (i.e., plural
noun); thus, the suggested word must also be a ``NNS". In this work,
we use \spacy, a free open-source library for \nlp, for extracting PoS
tags for each suggested word.

Second, it is required that the sentiment of the expanded \sent
becomes the same as the seed \sent. To ensure this, the suggested
words must be \neu.

Third, we additionally verify that the expanded \sents satisfied the
same search rules for the seed \sent in LC1 and LC2. This criteria
cannot be applied to other linguistic capabilities because they have
additional transformation templates.

In this work, out of the BERT suggested words, we randomly selects
words that matches their PoS tag of the expanded symbol in the parse
tree. The words that meets the second and third criteria are finally
employed for the expanded \sents.
% Our goal for generating the expanded \sents is to use them for
% evaluating the \sa models on the associated \lc in addition to the
% seed \sent. It is only achieved when the expanded \sents are also met
% with the same search rules for the \lc. For LC1 as an example, the
% expanded \sent must still conform to the specification of the seed's
% \lc specification. Therefore, the expanded \sents are required to be
% short and to only have \neu adjectives and \nns.
% \sw{Say why we use
%   the third criteria only for LC1 and LC2.}  \jl{I added the
%   explanation}

\paragraph{Running example.} The third step in Figure~\ref{fig:ExpEx} shows the words suggested
by BERT. For this masked \sent, BERT suggested six words. Each word is
associated with the confidence score provided by BERT, the PoS tag,
and the sentiment. Among the six words, only ``ways'' and ``things''
are validated by \tool{} because they have the Pos tag ``NNS'' and are
\neu. In addition, it is found that both \sents meets the search rule
of the associated \lc of \SareqExOne. In the end, two \sents of ``Or
both ways'' and ``Or both things'' are generated.

@@

% \subsubsection{\Sent Selection}
% \sw{@Jaeseong: Add how we select expanded \sents and motivate why.}

%% \paragraph{Running example.} \jl{I dont think we need this
%%   subsection of \sent selection it is already explained at the
%%   previous stage with running example.} \sw{This is to use the bert score to rank and select. Did we discuss this already?}
